Speed estimation for vehicles using \\SVM classification and edge detection

### 2016-06-15 Administratie/report/paper.tex
update report
kleine aanpassingen
Afwerking tekst

\subsection{SVM classification}\label{results-svm-classification}
We iterated through several parameter values (as discussed in \ref{parameters-classification}) to find optimal classification for road detection. For each combination of parameters we train and test on different combinations of datasets. For each combination we calculate the F1-scores \cite. These scores are used to compare the classification results (higher F-scores indicate better classification).

@@

\subsubsection{Block size}
In general, block sizes 16x16px and 32x32px tend to show the better results. As an example we present the F-scores for using color feature vectors with road marks in figure \ref{fig:col-wi-rm}. Other combinations follow the same trend, although 32x32px sometimes performs slightly better. For the goal of this application, a smaller block size will be more practical (i.e. better accuracy). This is why we prefer the block size of 16x16px.

@@

\subsubsection{LBP}
The use of LBP is crucial to our implementation in order to get a viable classifier for road detection. If we compare results (block size 16x16px) of feature vectors on LBP only to the combination of LBP and color, we get different results depending on whether or not we include road marks.

If we do not include road marks in the training data, we get slightly better results when combining LBP and color for the feature vectors, as shown in figure \ref{fig:16_wo_roadmarks}.
If on the other hand we do include these road marks, using only LBP shows equal or better results than the combination of LBP and color, as shown in figure \ref{fig:16_wi_roadmarks}.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{fig/16_wo_roadmarks.png}
\caption{Comparision of F-scores for 16x16px block size without road marks using RBF kernel }
\label{fig:16_wo_roadmarks}
\end{figure}

A possible explanation for this phenomenon is that when only asphalt is considered to be road, color is a much more consistent factor to weigh in to the classification. However when including road marks as part of the road, color intensity is no longer an added value and even tends to mislead the classifier.

Overall we can conclude that using only LBP with road marks shows the optimal results for our classification problem.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{fig/16_wi_roadmarks.png}
\caption{Comparision of F-scores for 16x16px block size with road marks using RBF kernel }
\label{fig:16_wi_roadmarks}
\end{figure}

@@

\textbf{TODO} check if list not f***ed up.\\
\noindent In retrospect, the optimal parameters for our classification problem are:
\begin{itemize}
\item RBF kernel
\item Block size 16x16px
\item Using only LBP histogram values for the feature vectors
\item Including road marks as part of the road  
\end{itemize}

@@

\begin{figure}[t]
\centering
\begin{subfigure}[t]{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/zebra_edge_filtered.png}
  \caption{Zebra crossing filtered out by edge detection.}
  \label{zebra-1}
\end{subfigure}%
\begin{subfigure}[t]{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/zebra_road_filtered.png}
  \caption{Zebra crossing classified as road.}
  \label{zebra-2}
\end{subfigure}
\caption{Zebra crossing\label{zebra}}
\end{figure}

\section{Conclusion}
\newpage
\bibliography{Computervisie}{}
\bibliographystyle{plain}

>>>>

\subsection{SVM classification}\label{results-svm-classification}
We iterated through several parameter values (as discussed in \ref{parameters-classification}) to find optimal classification for road detection. For each combination of parameters we train and test on different combinations of datasets. For each combination we calculate the F1-scores \cite{Powers2007}. These scores are used to compare the classification results (higher F-scores indicate better classification).

@@

\subsubsection{Block size}
In general, block sizes 16x16px and 32x32px tend to show the better results. As an example we present the F-scores for using color feature vectors with road marks in figure \ref{fig:col-wi-rm}. Other combinations follow the same trend, although 32x32px sometimes performs slightly better. This is why we conclude the block size of 32x32px to be a more reliable choice.

@@

\subsubsection{Feature vector values}
The use of LBP represents no significant gains in order to get a viable classifier for road detection. If we compare results (block size 32x32px) of feature vectors on color only to the combination of LBP and color,  performance remains roughly the same.
Therefore, in the case of our classification problem, we do not add LBP to the feature vector.

The best results are found when using color histogram values for the feature vector. This implies the combinations of color values are sufficient to distinguish the road from other surroundings. A possible explanation for this phenomenon is that when asphalt and road marks are considered to be road, color is a much more consistent factor to weigh in to the classification.

\textbf{TODO} grafiek wo/wi RM

\subsubsection{Road marks}
Including road marks as part of the road gets better results for the classification problem.
For our application this is more practical, because road marks should not be considered to be obstacles.

Overall we can conclude that using only color with road marks shows the optimal results for our classification problem.

TODO figure aanpassen
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{fig/16_wi_roadmarks.png}
\caption{Comparision of F-scores for 16x16px block size with road marks using RBF kernel }
\label{fig:16_wi_roadmarks}
\end{figure}

@@

\textbf{TODO} check if list not f***ed up.\\
\noindent In retrospect, the optimal parameters for our classification problem are:
\begin{itemize}
\item RBF kernel
\item Block size 32x32px
\item Using only color histogram values for the feature vectors
\item Including road marks as part of the road  
\end{itemize}

@@

\begin{figure}[H]
\centering
\begin{subfigure}[t]{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/zebra_edge_filtered.png}
  \caption{Zebra crossing filtered out by edge detection.}
  \label{zebra-1}
\end{subfigure}%
\begin{subfigure}[t]{.5\textwidth}
  \centering
  \includegraphics[width=.9\textwidth]{fig/zebra_road_filtered.png}
  \caption{Zebra crossing classified as road.}
  \label{zebra-2}
\end{subfigure}
\caption{Zebra crossing\label{zebra}}
\end{figure}

\section{Conclusion}
In this paper we combined SVM classification with edge detection in order to calculate the maximum speed a vehicle can drive, considering possible obstacles on the given trajectory.

The SVM classification shows optimal results when using color histogram values for the feature vector. These feature vectors are calculated for each block of 32x32px in an image.
These results tend to get better when we include road marks as part of the road in the training data. This is also optimal for use in our application as we don't want to detect road marks as obstacles.

To perform edge detection, the frame is first treated with morphological processing operations to remove insignificant details. In our application we used Canny Edge Detection to detect the side of the road and possible obstacles.

In order to combine the SVM classification with the edge detection, two strategies were proposed. The first strategy simply intersects the SVM classification results and edges to detect the nearest obstacle in the trajectory. This method show to be the safe option, by often falsely detecting an obstacle.  Results show that the best strategy is to first detect all the edges and then classify these edges as being part of the road or not, using the SVM classifier. Only edges which are not considered part of the road remain. The maximum speed is then calculated based on the nearest edge in the trajectory of the car.

\newpage
\bibliography{Computervisie}{}
\bibliographystyle{plain}
